---
title: "a3_task3_mariano_viz"
author: "Mariano Viz"
date: "01/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)

```



## Summary

In this report, I first compare purple and red urchins diameters. Then, I explore purple urchin counts by site and year.

## Citation

[Santa Barbara Coastal LTER and D. Reed. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative](https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b)


## Read in the data using `here()` and update the column names to lower snake case using `clear_names()`

```{r}

urchins_data <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()

```




# **Section 1**: Size comparisons, purple and red urchins

In this section, I compare purple and red urchins diameters using data visualization, descriptive statistics and performing a two-sample t-test.

### **Section 1.1**: Exploratory data visualization

Here, I compare the distribution of purple and red urchins diameters using a faceted histogram and a boxplot.

```{r}

ggplot(data = urchins_data, aes(x = size)) +
  geom_histogram(bins = 18) +
  facet_wrap(~common_name) +
  labs(x = "Urchin diameter (cm)",
       y = "Count",
       title = "Purple and Red Urchin diameter")

```



```{r}
ggplot(data = urchins_data, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(x = "Common name",
       y = "Urchin diameter (cm)",
       title = "Purple and Red Urchin diameter")

```

### **Section 1.2**: Summary Statistics

Here, I create a table with some descriptive statistics (mean, median, and standard deviation) of the purple and red urchins diameters. 


```{r}

urchins_stats <- urchins_data %>% 
  group_by(common_name) %>% 
  summarise(mean_size = mean(size),
            median_size = median(size),
            sd_size = sd(size),
            n = n()) %>%
  kbl()%>% 
  kable_styling()

urchins_stats

```

### **Section 1.3**: Size comparisons, purple and red urchins

Here, I describe the observed differences between purple and red urchins diameters. To help contextualize,  I perform a two-sample t-test and I create a table with the mean values, mean difference, and mean percentage difference between purple and red urchins diameters.

```{r}
urchins_means <- urchins_data %>% 
  group_by(common_name) %>% 
  summarise(mean_size = mean(size)) %>% 
  pivot_wider(names_from = common_name,
              values_from = mean_size) %>% 
  clean_names() %>% 
  mutate(mean_diff = red_urchin - purple_urchin) %>% 
  mutate(mean_diff_perc = mean_diff / red_urchin *100) %>% 
  kbl()%>% 
  kable_styling()
  

urchins_means

purple_sample <- urchins_data %>% 
  filter(common_name == "Purple Urchin") %>% 
  pull(size)

red_sample <- urchins_data %>% 
  filter(common_name == "Red Urchin") %>% 
  pull(size)

t.test(x = purple_sample, red_sample)


```

From the data visualization we can see that observations are normally distributed. At the same time, because we have a really big sample size (way over 30) we can be confident that the distribution of means is going to be normal (Central Limit Theorem). 
There is a marked difference between the means of purple and red urchins diameters (red urchins mean diameter is 34% higher than that of purple urchins) and the probability of having taken this samples from populations with the same mean is extremely low (p-value < 2.2e-16).   


# **Section 2**: Exploration of purple urchin counts by site and year

In this section, I explore purple urchin counts by site and year.

### **Section 2.1**: Data Wrangling

Here, I create a subset of the urchins data that only includes observations for purple urchins. I also convert the date to class "Date" and create new columns containing the year and full name of the sites. 


```{r}

purple_urchins <- urchins_data %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(site_full = case_when(site == "NAPL" ~ "Naples Reef",
                               site == "AQUE" ~ "Arroyo Quemado Reef",
                               site == "CARP" ~ "Carpinteria Reef",
                               site == "MOHK" ~ "Mohawk Reef",
                               site == "IVEE" ~ "Isla Vista Reef"))

```


### **Section 2.2**: Total number of purple urchins observed by site and year

Here, I use `count()` to get the number of purple urchins observed by site and year and then I make an exploratory graph of the counts of purple urchins recorded at each site over time.

```{r}

purple_stats <- purple_urchins %>% 
  count(site_full, year)


ggplot(data = purple_stats, aes(x = year, y = n)) +
  geom_col() +
  facet_wrap(~site_full) +
  labs(x = "Year",
       y = "Observations",
       title = "Total number of Purple Urchins observed by site and year")


```

